Comparing OTU quality across clustering strategies

### 2021-06-08 exploratory/2021/compare-strategies_2021-05.Rmd
Reformat strategy & method labels gotta have 'de novo' italicized

opticlust <- read_tsv(here('subworkflows/1_prep_samples/results/opticlust_results.tsv')) %>% 
  full_join(read_tsv(here('subworkflows/1_prep_samples/results/dataset_sizes.tsv'))) %>% 
  mutate_perf() %>% mutate(fraction_mapped = 1, strategy = method)
sum_opticlust <- opticlust %>% group_sum()
optifit_dbs <- read_tsv(here('subworkflows/2_fit_reference_db/results/optifit_dbs_results.tsv')) %>% 
  mutate_perf()
optifit_split <- read_tsv(here('subworkflows/3_fit_sample_split/results/optifit_split_results.tsv')) %>% 
  mutate_perf()
optifit_all <- list(optifit_dbs %>% 
                   mutate(strategy = glue('database_{ref}')),
                 optifit_split %>% 
                   mutate(strategy = 'self-split')) %>% 
  reduce(full_join)
sum_optifit <- optifit_all %>% group_sum()
vsearch <- read_tsv(here('subworkflows/4_vsearch/results/vsearch_results.tsv')) %>% 
  mutate_perf() %>% 
  mutate(strategy = case_when(
    method == 'de_novo' ~ method,
    TRUE ~ as.character(glue('database_{ref}')))) 
mothur_vsearch <- list(optifit_all, opticlust, vsearch) %>% 
  lapply(select_cols) %>% 
  reduce(bind_rows) %>% 
  mutate(strategy = factor(strategy, levels = c('database_rdp', 'database_silva', 'database_gg', 'self-split', 'de_novo')),
         method = factor(method, levels = c('open', 'closed', 'de_novo')))
sum_all <- mothur_vsearch %>% group_sum()
head(sum_all)
```

@@

- [x] vertical jitter
- [x] frac map only makes sense for closed
- [x] dotted line for _de novo_
- [x] filled/unfilled for closed/open
- [x] box around facets
- [x] bigger points
- [ ] set theme elements to add space between facets: `+ theme(panel.spacing = unit(c(3,3),'lines'))`
- generally: build for the dimensions required by the journal

@@

```{r otu-quality_mothur-vs-vsearch, fig.width = 6, fig.height = 4.5}
plot_quality <- function(dat, y_val, title = '') {
  dat %>% 
    ggplot(aes(strategy, {{ y_val }}, 
               color = tool, 
               shape = method)) + 
    geom_point(size = 3, position = position_dodge(width = 0.4)) +
    facet_wrap(dataset ~ ., nrow=1) +
    scale_shape_manual(values = list(open = 1, closed = 19, de_novo = 17)) +
    #scale_color_manual(values = tri_colors) +
    scale_y_continuous(labels = c('0', '0.5', '1'), 
                       breaks = c(0, 0.5, 1),
                       limits = c(0, 1)) +
    coord_flip() +
    labs(x = '', y = '', title = title) +
    theme_bw() +
    theme(legend.position="none")
}

@@

shared_legend <- get_legend(mcc_plot + 
                              guides(color = guide_legend(nrow = 1)) +
                              theme(legend.position = "bottom",
                                    legend.title = element_blank())
                            )

@@

```{r runtime_mothur-vs-vsearch}
plot_runtime <- function(dat, yval, title = '') {
  dat %>% 
    ggplot(aes(strategy, {{ yval }}, 
               color = tool, 
               shape = method
               )) + 
    geom_point(size = 3, position = position_dodge(width = 0.4)) +
    facet_wrap(dataset ~ ., nrow = 1, scales = 'free_x') +
    scale_shape_manual(values = list(open = 1, closed = 19, de_novo = 17)) +
    scale_y_log10() +
    #scale_color_manual(values = tri_colors) +
    coord_flip() +
    labs(x = '', y = '', title = title) +
    theme_bw() +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "none")
}
runtime_plot <- sum_all %>% plot_runtime(sec_median) + labs(title = 'Runtime (sec)')
runtime_plot
```

>>>>

opticlust <- read_tsv(here('subworkflows/1_prep_samples/results/opticlust_results.tsv')) %>% 
  full_join(read_tsv(here('subworkflows/1_prep_samples/results/dataset_sizes.tsv'))) %>% 
  mutate_perf() %>% mutate(fraction_mapped = 1, strategy = method)
sum_opticlust <- opticlust %>% group_sum()
optifit_dbs <- read_tsv(here('subworkflows/2_fit_reference_db/results/optifit_dbs_results.tsv')) %>% 
  mutate_perf()
optifit_split <- read_tsv(here('subworkflows/3_fit_sample_split/results/optifit_split_results.tsv')) %>% 
  mutate_perf()
optifit_all <- list(optifit_dbs %>% 
                   mutate(strategy = glue('database_{ref}')),
                 optifit_split %>% 
                   mutate(strategy = 'self-split')) %>% 
  reduce(full_join)
sum_optifit <- optifit_all %>% group_sum()
vsearch <- read_tsv(here('subworkflows/4_vsearch/results/vsearch_results.tsv')) %>% 
  mutate_perf() %>% 
  mutate(strategy = case_when(
    method == 'de_novo' ~ method,
    TRUE ~ as.character(glue('database_{ref}')))) 
mothur_vsearch <- list(optifit_all, opticlust, vsearch) %>% 
  lapply(select_cols) %>% 
  reduce(bind_rows) %>% 
  mutate(method = as.character(method),
         strategy = as.character(strategy))
sum_all <- mothur_vsearch %>% group_sum()
head(sum_all)
```

### make the strategy & method labels prettier

```{r format-labels}
sum_all <- sum_all %>%
  mutate(
    strategy = factor(
      case_when(
        strategy == "de_novo"        ~ "_de novo_",
        strategy == 'database_rdp'   ~ "db: RDP",
        strategy == 'database_silva' ~ "db: SILVA",
        strategy == 'database_gg'    ~ "db: Greengenes",
        TRUE                         ~ strategy
      ),
      levels = c('db: RDP', 'db: SILVA', 'db: Greengenes', 
                 'self-split',  '_de novo_')
    ),
    method = factor(
      case_when(method == "de_novo" ~ "_de novo_",
                TRUE                ~ method),
      levels = c('open', 'closed', '_de novo_')
    )
  )
head(sum_all)
```

@@

- [x] vertical jitter
- [x] frac map only makes sense for closed
- [x] dotted line for _de novo_
- [x] filled/unfilled for closed/open
- [x] box around facets
- [x] bigger points
- [ ] set theme elements to add space between facets: `+ theme(panel.spacing = unit(c(3,3),'lines'))`
- generally: build for the dimensions required by the journal
- [ ] fix strategy labels (remove underscores, italicize _de novo_)

@@

```{r otu-quality_mothur-vs-vsearch, fig.width = 6, fig.height = 4.5}
plot_quality <- function(dat, y_val, title = '') {
  dat %>% 
    ggplot(aes(strategy, {{ y_val }}, 
               color = tool, 
               shape = method)) + 
    geom_point(size = 3, position = position_dodge(width = 0.4)) +
    facet_wrap(dataset ~ ., nrow=1) +
    scale_shape_manual(values = list(open = 1, closed = 19, `_de novo_` = 17)) +
    #scale_color_manual(values = tri_colors) +
    scale_y_continuous(labels = c('0', '0.5', '1'), 
                       breaks = c(0, 0.5, 1),
                       limits = c(0, 1)) +
    coord_flip() +
    labs(x = '', y = '', title = title) +
    theme_bw() +
    theme(legend.position="none",
          axis.text.y = element_markdown())
}

@@

shared_legend <- get_legend(mcc_plot +
                              guides(color = guide_legend(nrow = 1)) +
                              theme(legend.position = "bottom",
                                    legend.title = element_blank(),
                                    legend.text = element_markdown())
                            )

@@

```{r runtime_mothur-vs-vsearch}
plot_runtime <- function(dat, yval, title = '') {
  dat %>% 
    ggplot(aes(strategy, {{ yval }}, 
               color = tool, 
               shape = method
               )) + 
    geom_point(size = 3, position = position_dodge(width = 0.4)) +
    facet_wrap(dataset ~ ., nrow = 1, scales = 'free_x') +
    scale_shape_manual(values = list(open = 1, closed = 19, `_de novo_` = 17)) +
    scale_y_log10() +
    #scale_color_manual(values = tri_colors) +
    coord_flip() +
    labs(x = '', y = '', title = title) +
    theme_bw() +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "none",
          axis.text.y = element_markdown())
}
runtime_plot <- sum_all %>% plot_runtime(sec_median) + 
  labs(title = 'Runtime (sec)')
runtime_plot
```
